---
title: "practical_exercise_10 , Methods 3, 2021, autumn semester"
author: 'Anja, Astrid, Jessica, Juli and Magnus'
date: "15/11/21"
output: html_document
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>

```{r, include = FALSE}
library(reticulate)
options(reticulate.repl.quiet = TRUE) 
```

```{python}
import numpy as np
import matplotlib.pyplot as plt
import seaborn as sns

from sklearn.preprocessing import StandardScaler
scaler = StandardScaler() 

from sklearn.model_selection import StratifiedKFold
StratifiedKFold = StratifiedKFold()

from sklearn.model_selection import cross_val_score
cross_val_score = cross_val_score()

np.random.seed(7)
```

```{python, include = FALSE}
def equalize_targets(data, y):
    np.random.seed(7)
    targets = np.unique(y)
    counts = list()
    indices = list()
    for target in targets:
        counts.append(np.sum(y == target))
        indices.append(np.where(y == target)[0])
    min_count = np.min(counts)
    first_choice = np.random.choice(indices[0], size=min_count, replace=False)
    second_choice = np.random.choice(indices[1], size=min_count, replace=False)
    third_choice = np.random.choice(indices[2], size=min_count, replace=False)
    fourth_choice = np.random.choice(indices[3], size=min_count, replace=False)
    
    new_indices = np.concatenate((first_choice, second_choice,
                                 third_choice, fourth_choice))
    new_y = y[new_indices]
    new_data = data[new_indices, :, :]
    
    return new_data, new_y
```

# Exercises and objectives

1) Use principal component analysis to improve the classification of subjective experience  
2) Use logistic regression with cross-validation to find the optimal number of principal components  


REMEMBER: In your report, make sure to include code that can reproduce the answers requested in the exercises below (__MAKE A KNITTED VERSION__)  
REMEMBER: This is Assignment 4 and will be part of your final portfolio   

# EXERCISE 1 - Use principal component analysis to improve the classification of subjective experience  

We will use the same files as we did in Assignment 3
The files `megmag_data.npy` and `pas_vector.npy` can be downloaded here (http://laumollerandersen.org/data_methods_3/megmag_data.npy) and here (http://laumollerandersen.org/data_methods_3/pas_vector.npy)  
The function `equalize_targets` is supplied - this time, we will only work with an equalized data set. One motivation for this is that we have a well-defined chance level that we can compare against. Furthermore, we will look at a single time point to decrease the dimensionality of the problem  

##1.1) 
Create a covariance matrix, find the eigenvectors and the eigenvalues
###1.1.i. 
Load `megmag_data.npy` and call it `data` using `np.load`. You can use `join`, which can be imported from `os.path`, to create paths from different string segments 
```{python}
data = np.load("data/megmag_data.npy")
y = np.load("data/pas_vector.npy")
```
###1.1.ii. 
Equalize the number of targets in `y` and `data` using `equalize_targets`
```{python}
data_equal, y_equal = equalize_targets(data, y)
# lets have a lookie
data.shape
y.shape
data_equal.shape
y_equal.shape
```
###1.1.iii. 
Construct `times=np.arange(-200, 804, 4)` and find the index corresponding to 248 ms - then reduce the dimensionality of `data` from three to two dimensions by only choosing the time index corresponding to 248 ms (248 ms was where we found the maximal average response in Assignment 3) 
```{python}
times = np.arange(-200, 804, 4)
max_index = np.where(times == 248)[0][0] #The [0] thingeys are to get into the nr. not just the wierd arrays
max_index
data_248 = data_equal[:,:,max_index]
data_248.shape
```
>the equal data or the normal data? Cuz the times thingy wouldn't work for the equal data right?
>expect it will! cuz look at the time dim, that didn't get effected by the equalising

###1.1.iv. 
Scale the data using `StandardScaler` 
```{python}
data_248_scaled = scaler.fit_transform(data_248)
```

###1.1.v. 
Calculate the sample covariance matrix for the sensors (you can use `np.cov`) and plot it (either using `plt.imshow` or `sns.heatmap` (`import seaborn as sns`))
```{python}
cov_248 = np.cov(data_248_scaled, rowvar = False)
abs_cov_248 = np.abs(cov_248) # This is just to make it obvious that light is "bad" and dark is "good". Whether the covariance is not interesting right now.
plt.figure()
sns.heatmap(abs_cov_248, square = True)
plt.show()
```

###1.1.vi. 
What does the off-diagonal activation imply about the independence of the signals measured by the 102 sensors?  
>The covariance is similar in big chucks - this might indicate sensors that are close spatially, but not in index. Thus they pick up some of the same activation. In data like this covariance can be seen because sensors from opposite brain hemispheres can pick up similar data.
Ideally we would have 0 cov exept for the diagonal. We could probably reduce the dimensions when we see so much covariance. 

###1.1.vii. 
Run `np.linalg.matrix_rank` on the covariance matrix - what integer value do you get? (we'll use this later)  
```{python}
np.linalg.matrix_rank(cov_248)
```
>97

###1.1.viii. 
Find the eigenvalues and eigenvectors of the covariance matrix using `np.linalg.eig` - note that some of the numbers returned are complex numbers, consisting of a real and an imaginary part (they have a _j_ next to them). We are going to ignore this by only looking at the real parts of the eigenvectors and -values. Use `np.real` to retrieve only the real parts  
```{python}
eigenvalues, eigenvectors = np.linalg.eig(cov_248)
sum(np.iscomplex(eigenvalues)) # Finding and counting the True values, that represent complex numbers
sum(np.iscomplex(eigenvectors))
# No complex numbers found, no np.real needed
```

##1.2) 
Create the weighting matrix $W$ and the projected data, $Z$

###1.2.i. 
We need to sort the eigenvectors and eigenvalues according to the absolute values of the eigenvalues (use `np.abs` on the eigenvalues).  
```{python}
eigenvalues = np.abs(eigenvalues)
eigenvectors = np.abs(eigenvectors)
```

###1.2.ii. 
Then, we will find the correct ordering of the indices and create an array, e.g. `sorted_indices` that contains these indices. We want to sort the values from highest to lowest. For that, use `np.argsort`, which will find the indices that correspond to sorting the values from lowest to highest. Subsequently, use `np.flip`, which will reverse the order of the indices.  
```{python}
sorted_indices = np.argsort(eigenvalues) #This array has the indices of the eigenvalues - but sorted so the index with the smallest value is first.
sorted_indices = np.flip(sorted_indices) # Now it is higest to lowest
#eigenvalues tells the length of the eigenvectors that have been normalised so that they are all have length 1.

```

###1.2.iii. 
Finally, create arrays of sorted eigenvalues and eigenvectors using the `sorted_indices` array just created. For the eigenvalues, it should like this `eigenvalues = eigenvalues[sorted_indices]` and for the eigenvectors: `eigenvectors = eigenvectors[:, sorted_indices]`
```{python}
eigenvalues
eigenvalues = eigenvalues[sorted_indices]
eigenvalues
eigenvectors = eigenvectors[:, sorted_indices]
```


```{python}
# Just simple code to understand how it works. Exept is doesn't?
a = np.arange(0, 11, 1)
a
a = np.random.shuffle(a)
a
sorted_indices = np.argsort(a)
sorted_indices
a[sorted_indices]
```

###1.2.iv. 
Plot the log, `np.log`, of the eigenvalues, `plt.plot(np.log(eigenvalues), 'o')` - are there some values that stand out from the rest? In fact, 5 (noise) dimensions have already been projected out of the data - how does that relate to the matrix rank (Exercise 1.1.vii)  
```{python}
plt.figure()
plt.plot(np.log(eigenvalues), 'o', markersize = 2)
plt.show()
```
>97 + 5 = 102

###1.2.v. 
Create the weighting matrix, `W` (it is the sorted eigenvectors)
```{python}
W = eigenvectors
```

###1.2.vi. 
Create the projected data, `Z`, $Z = XW$ - (you can check you did everything right by checking whether the $X$ you get from $X = ZW^T$ is equal to your original $X$, `np.isclose` may be of help)
```{python}
# Is this correct? (see slides week 10, pg. 7):
# X = n * d 

data_248_scaled.shape # is (396,102) so we have 396*102 (40392) observations. Which is n?
np.ndim(data_248) # is 2, which is d?

X = 40392 * 2
```


```{python}
# Or maybe it is just the data?
X = data_248_scaled
```


```{python}
Z = X @ W

#Checking that it is correct
X_2 = Z @ W.T
np.isclose(X, X_2) # It is not


```


###1.2.vii. 
Create a new covariance matrix of the principal components (n=102) - plot it! What has happened off-diagonal and why?
```{r}
# idk what prinicpal competent issss
```


# EXERCISE 2 - Use logistic regression with cross-validation to find the optimal number of principal components  

#2.1) 
We are going to run logistic regression with in-sample validation 
###2.1.i. 
First, run standard logistic regression (no regularization) based on ~~$Z_{d \times k}$~~ $Z_{n \times k}$ and `y` (the target vector). Fit (`.fit`) 102 models based on: $k = [1, 2, ..., 101, 102]$ and $d = 102$. For each fit get the classification accuracy, (`.score`), when applied to $Z_{d \times k}$ and $y$. This is an in-sample validation. Use the solver `newton-cg` if the default solver doesn't converge
###2.1.ii. 
Make a plot with the number of principal components on the _x_-axis and classification accuracy on the _y_-axis - what is the general trend and why is this so?
###2.1.iii. 
In terms of classification accuracy, what is the effect of adding the five last components? Why do you think this is so?
#2.2) 
Now, we are going to use cross-validation - we are using `cross_val_score` and `StratifiedKFold` from `sklearn.model_selection`
###2.2.i. 
Define the variable: `cv = StratifiedKFold()` and run `cross_val_score` (remember to set the `cv` argument to your created `cv` variable). Use the same `estimator` in `cross_val_score` as in Exercise 2.1.i. Find the mean score over the 5 folds (the default of `StratifiedKFold`) for each $k$, $k = [1, 2, ..., 101, 102]$  
###2.2.ii. 
Make a plot with the number of principal components on the _x_-axis and classification accuracy on the _y_-axis - how is this plot different from the one in Exercise 2.1.ii?
###2.2.iii. 
What is the number of principal components, $k_{max\_accuracy}$, that results in the greatest classification accuracy when cross-validated?  
###2.2.iv. 
How many percentage points is the classification accuracy increased with relative to the to the full-dimensional, $d$, dataset 
###2.2.v. 
How do the analyses in Exercises 2.1 and 2.2 differ from one another? Make sure to comment on the differences in optimization criteria.  
#2.3) 
We now make the assumption that $k_{max\_accuracy}$ is representative for each time sample (we only tested for 248 ms). We will use the PCA implementation from _scikit-learn_, i.e. import `PCA` from `sklearn.decomposition`.
###2.3.i. 
For __each__ of the 251 time samples, use the same estimator and cross-validation as in Exercises 2.1.i and 2.2.i. Run two analyses - one where you reduce the dimensionality to $k_{max\_accuracy}$ dimensions using `PCA` and one where you use the full data. Remember to scale the data (for now, ignore if you get some convergence warnings - you can try to increase the number of iterations, but this is not obligatory)  
###2.3.ii. 
Plot the classification accuracies for each time sample for the analysis with PCA and for the one without in the same plot. Have time (ms) on the _x_-axis and classification accuracy on the _y_-axis 
###2.3.iii. 
Describe the differences between the two analyses - focus on the time interval between 0 ms and 400 ms - describe in your own words why the logistic regression performs better on the PCA-reduced dataset around the peak magnetic activity  

